Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-10-13 Manuscript/manuscript.tex
Small updates to the text in Stoichiometry section.

In addition, we estimated relative binding affinities 
of several moieties in POPC towards \ce{Ca^{2+}}. 
The dominant contribution
to the binding of \ce{Ca^{2+}} to POPC membranes comes from the phosphate group. 
Such a finding is based on the calculated
number of contacts per lipid between \ce{Ca^{2+}} and POPC oxygen atoms, 
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O}$,
in which a contact between the two atoms is counted 
if their distance is less than or equal $0.3 \, \mathrm{nm}$ 
(robustness of this definition discussed in reference~\citenum{javanainen17}). 
The average number of contacts per lipid between \ce{Ca^{2+}} and any oxygen atom in POPC is 
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 $,    % = 33.8/128 
whereas if only phosphate oxygen atoms are considered 
the average number of contacts decreases only by a tiny amount to 
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.262 $,    % = 33.5/128 . 
and if only carbonyl oxygen atoms are considered 
(phosphate oxygens contribute, but are not counted as contacts),
this value is merely
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.066 $.    % = 8.4/128 . 
This means that calcium is in contact with a phosphate oxygen for 99\% of the time it is bound to the membrane, 
whereas only 25\% of the time accounts for contact with a carbonyl oxygen,
and purely carbonyl bound states account for less than 1\% of configurations. 
We corroborate this finding with the probability isodensity contours in Fig.~\ref{fig:volmaps}, 
which are easier to interpret.

@@

Internal side-note: 
However, $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$ is increased 
when carbonyl dipoles are scaled only with 0.945 
(roughly done through charge redistribution)
 to : $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} \approx 0.08 $.    % = (70.-59.8)/128 .

>>>>

In addition, we estimated relative binding affinities 
of several moieties in POPC towards \ce{Ca^{2+}}. 
The dominant contribution
to the binding of \ce{Ca^{2+}} to POPC membranes comes from the phosphate group. 
Such a finding is based on the calculated
number of contacts per lipid between \ce{Ca^{2+}} and POPC oxygen atoms, 
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O}$,
in which a contact between the two atoms is counted 
if their distance is less than or equal $0.3 \, \mathrm{nm}$ 
(robustness of this definition discussed in reference~\citenum{javanainen17}). 
% Numbers from simulation with 346mM conc. CaCl2
The average number of contacts per lipid between \ce{Ca^{2+}} and any oxygen atom in POPC is 
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.134 $,    % = 17.1/128 
whereas if only phosphate oxygen atoms are considered 
the average number of contacts decreases only by a tiny amount to 
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.133 $,    % = 17.0/128 . 
and if only carbonyl oxygen atoms are considered 
(phosphate oxygens contribute, but are not counted as contacts),
this value is merely
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.033 $.    % = 4.3/128 . 
% Numbers from simulation with 980/869mM conc. CaCl2
%The average number of contacts per lipid between \ce{Ca^{2+}} and any oxygen atom in POPC is 
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 $,    % = 33.8/128 
%whereas if only phosphate oxygen atoms are considered 
%the average number of contacts decreases only by a tiny amount to 
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.262 $,    % = 33.5/128 . 
%and if only carbonyl oxygen atoms are considered 
%(phosphate oxygens contribute, but are not counted as contacts),
%this value is merely
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.066 $.    % = 8.4/128 . 
This means that calcium is in contact with a phosphate oxygen for 99\% of the time it is bound to the membrane, 
whereas only 25\% of the time accounts for contact with a carbonyl oxygen,
and purely carbonyl bound states account for less than 1\% of configurations. 
We corroborate this finding with the probability isodensity contours in Fig.~\ref{fig:volmaps}, 
which are easier to interpret.

@@

Internal side-note: 
However, $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$ is increased 
in simulation withy the highest \ce{CaCl2} concentration
when carbonyl dipoles are scaled only with 0.945 
(roughly done through charge redistribution)
 to : $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} \approx 0.08 $.    % = (70.-59.8)/128 .
